---
title: "Create synthetic data"
author: "Celeste Falcon [email: celestemfalcon@gmail.com]"
date: today
format: pdf
editor: visual
editor_options: 
  chunk_output_type: console
---

```{r packages and functions}
# Install the synthpop package
# install.packages("synthpop")

# Load the library
library(synthpop)
library(readxl)
library(tidyverse)
library(broman)

```

```{r example}
# Load your original data (replace 'original_data' with the name of your data frame)
# Example with a built-in R dataset:
data(iris) 
original_data <- iris

# Generate the synthetic data set
# The syn() function automatically determines the best distribution for each variable
synthetic_data <- syn(original_data)

# Extract the synthetic data frame
synthetic_df <- synthetic_data$syn

# Compare the statistical properties of the original and synthetic data (optional)
# The package provides tools for this, such as the compare.syn function
compare(synthetic_data, original_data)
```

```{r input data}
bioassay_data <- read_xlsx("./input/anonymized_for_portfolio/Bioassay Results_anonymized.xlsx", sheet = 5) %>%
  filter(!is.na(Dead)) %>% 
  mutate(Week = as.factor(Week)) %>% 
  mutate(Sample = as.factor(Sample)) %>% 
  mutate(Replicate = as.factor(Replicate)) %>% 
  mutate(Duration = as.factor(Duration))
```

```{r split into control and treatment datasets}

bioassay_data_control <- bioassay_data %>% 
  filter(Sample == "CONTROL")
  
bioassay_data_treatment <- bioassay_data %>% 
  filter(Sample == "TREATMENT")
```

```{r generate synth data}
#| echo: true
# # full data set
# bioassay_data_syn <- syn(bioassay_data, visit.sequence = c("Dead", "Total", "Mortality"))
# bioassay_data_syn_df <- bioassay_data_syn$syn
# compare(bioassay_data_syn, bioassay_data, vars = c("Mortality"), plot = F)

# control data set
bioassay_data_control_syn <- syn(bioassay_data_control, visit.sequence = c("Dead", "Total", "Mortality"))
bioassay_data_control_syn_df <- bioassay_data_control_syn$syn
# compare(bioassay_data_control_syn, bioassay_data_control, vars = c("Mortality"), plot = T)

compare(bioassay_data_control_syn, bioassay_data_control, vars = c("Mortality"), plot = F)

# treatment data set
bioassay_data_treatment_syn <- syn(bioassay_data_treatment, visit.sequence = c("Dead", "Total", "Mortality"))
bioassay_data_treatment_syn_df <- bioassay_data_treatment_syn$syn
compare(bioassay_data_treatment_syn, bioassay_data_treatment, vars = c("Mortality"), plot = F)

```

Distribution plots not shown to protect client's real data.

pMSE is propensity mean squared error, a measure of the utility of synthetic data based on how distinguishable it is from the original, real data. Values closer to 0 indicate high utility (the synthetic data is indistinguishable from the real data).

S_pMSE is its standardized ratio to account for scale of measurement for each variable

```{r output synth data}
# write.csv(bioassay_data_syn_df, "./synthetic_data.csv", row.names = F)

bioassay_data_ctrl_trt_syn <- bind_rows(bioassay_data_control_syn_df, bioassay_data_treatment_syn_df)

# write.csv(bioassay_data_ctrl_trt_syn, "./synthetic_data_by_sample.csv", row.names = F)

```
