---
title: "Analysis of Mortality on Anonymized, Simulated Data"
subtitle: "Simulated within sample"
author: "Celeste Falcon [email: celestemfalcon@gmail.com]"
date: today
format: pdf
editor: visual
editor_options: 
  chunk_output_type: console
---

```{r packages and functions}
#| include: false
#| 
library(readxl)
library(tidyverse)
library(broman)
library(knitr)
library(fmsb)
library(rstatix)
```

```{r input data}
#| include: false
#| 
bioassay_data <- read.csv("./synthetic_data_by_sample.csv") %>%
  filter(!is.na(Dead)) %>% 
  mutate(Alive = Total - Dead) %>% 
  mutate(`%Mortality` = as.numeric(myround(Mortality * 100)))
```

```{r check input data against original}
#| include: false
#| 
bioassay_data_means <- bioassay_data %>% 
  group_by(Week, Sample, Duration) %>% 
  mutate(Mean_mortality = mean(`%Mortality`, na.rm = T))
```

# Experimental Design and Analysis

*Anopheles quadrimaculatus* third instar larvae were exposed to **TREATMENT** larvicide. Mortality was assessed at 24, 48, 72, and 96 hours, with 4 replicates each (n = 25 larvae per replicate). This report contains assay results for a simulated dataset.

## Analytic Methods

All analyses were completed in R ver. 4.5.2 [^1]. All comparisons were performed with a Fisher’s Exact Test, as expected cell means were \< 5 [^2]. For all statistical tests, significance was assessed at alpha = 0.05 (i.e., p-value \< or = 0.05 demonstrated a significant difference).

[^1]: R Core Team (2025). *R: A Language and Environment for Statistical Computing*. R Foundation for Statistical Computing, Vienna, Austria. <https://www.R-project.org/>.

[^2]: Zar, J. H. 2009. Biostatistical Analysis (5th ed.). Prentice Hall: Upper Saddle River, NJ, USA.

### Mortality of control replicates

Mortality within control replicates was calculated. Mortality of controls should be \< 10%. Replicates with mortality = or \> 10% were identified as outliers.

### Analysis of treatment replicates

A two-sided Fisher’s Exact Test was used to perform pairwise comparisons of mortality for treatment replicates. Replicates that were significantly different from others were identified as outliers. The Holm method was used to adjust p-values for multiple comparisons. In situations where two of the replicates were significantly different from the remaining two replicates, the replicates with higher mortality values were treated as outliers.

### Comparison of treatment to control

A one-sided Fisher’s Exact Test was used to determine if mortality for the treatment group was greater than the control. When outliers were detected in either control or treatment replicates, the test was repeated excluding the outliers.

{{< pagebreak >}}

## Synopsis

### Mortality of control replicates

Mortality of control replicates was \< 10% for all durations considered in week 1. Some replicates from weeks 2, 3, 4, 5, and 6 had mortality \> or = 10%.

### Analysis of treatment replicates

There were no significant differences among treatment replicates for weeks 2, 5, and 6. There were significant differences among replicates for week 1 (replicate 1 at 48 hours treated as an outlier), week 3 (replicate 4 at 48 hours and replicate 1 at 96 hours treated as outliers), and week 4 (replicate 2 at 48 hours treated as an outlier).

### Comparison of treatment to control

For all weeks, mortality in the treatment group was significantly greater than mortality in the control group, regardless of the presence of outliers in the dataset.

{{< pagebreak >}}

## Week 1 results

For week 1, there were four replicates (1, 2, 3, 4) for both the control and treatment groups. Mortality was assessed at 24 and 48 hours.

### Mortality of control replicates

```{r}
#| include: false
#| 
bioassay_data %>% 
  filter(Week == "1") %>% 
  filter(Sample == "CONTROL") %>% 
  mutate(Outlier = if_else(`%Mortality` >= 10, "Yes", "No"))

bioassay_data_control_outliers <-  bioassay_data %>% 
  filter(Sample == "CONTROL") %>% 
  mutate(Outlier = if_else(`%Mortality` >= 10, "Yes", "No"))
```

For all control replicates in week 1, mortality was \< 10%.

```{r}
#| include: false
kable(bioassay_data %>% 
        filter(Week == "1") %>% 
        filter(Sample == "CONTROL") %>% 
        mutate(Outlier = if_else(`%Mortality` >= 10, "Yes", "No")) %>% 
        select(Duration, Replicate, `%Mortality`, Outlier))
```

### Analysis of treatment replicates

```{r}
#| echo: false
#| 
week <- unique(bioassay_data$Week)
duration <- unique(bioassay_data$Duration)
replicate <- unique(bioassay_data$Replicate)

for (i in week) {
  for (j in duration) {
    for (k in replicate) {
      # i = week[1]
      # j = duration[1]
      # k = replicate[1]
      
      data_treatment_weeki_durationj_repk <- bioassay_data %>% 
        filter(Sample == "TREATMENT") %>% 
        filter(Week == i) %>% 
        filter(Duration == j) %>% 
        filter(Replicate == k)
      
      assign(paste("data_treatment_week", i, "duration", j, "rep", replicate[k], sep = "_"), data_treatment_weeki_durationj_repk)
      
    }
  }
}
```

```{r compare reps pairwise}
#| echo: false
#| 
for (i in week) {
  for (j in duration) {
    
    # i = week[1]
    # j = duration[1]
    
    #### using rstatix function ####
    
    dead_and_alive <- bioassay_data %>% 
      filter(Sample == "TREATMENT") %>% 
      filter(Week == i) %>% 
      filter(Duration == j) %>% 
      select(Dead, Alive)  %>% 
      as.data.frame()
    
    if (nrow(dead_and_alive) > 0) {
      rownames(dead_and_alive) <- c("rep 1", "rep 2", "rep 3", "rep 4")
      
      fisher.test(dead_and_alive)
      compare_all_trt_reps_pvalue <- fisher.test(dead_and_alive)$p.value
      
      assign(paste("compare_all_trt_reps_pvalue_week", i, "duration", j, sep = "_"), compare_all_trt_reps_pvalue)
      
      
      pairwise_fisher_test_results <- pairwise_fisher_test(dead_and_alive, p.adjust.method = "holm")
      
      assign(paste("pairwise_fisher_test_results_week", i, "duration", j, sep = "_"), pairwise_fisher_test_results)
      
    }
  }
}
```

During week 1, readings taken at 48 hours had significantly different mortality between replicates.

**Readings at 48 hours:** Replicate 1 (68% mortality) was significantly different from replicates 2 (100% mortality) and 4 (100% mortality). Replicate 1 was treated as an outlier.

```{r}
#| include: false
compare_all_trt_reps_pvalue_week_1_duration_24hr
compare_all_trt_reps_pvalue_week_1_duration_48hr

kable(pairwise_fisher_test_results_week_1_duration_24hr)
kable(pairwise_fisher_test_results_week_1_duration_48hr)
```

```{r}
#| include: false
bioassay_data_treatment_outliers_week_1 <- bioassay_data %>% 
  filter(Sample == "TREATMENT") %>% 
  filter(Week == 1) %>% 
  mutate(Outlier = if_else(Duration %in% c("48hr") & Replicate == "1", "Yes", "No"))

kable(bioassay_data_treatment_outliers_week_1)
```

| Duration | Replicates for comparison | Adjusted p-value |
|----------|---------------------------|-----------------:|
| 48hr     | 1, 2                      |           0.0242 |
| 48hr     | 1, 4                      |           0.0242 |

### Comparison of treatment to control

Mortality was significantly greater in the treatment than in the control for each experiment duration (24 and 48 hours) in week 1.

```{r}
#| echo: false
#| 
for (i in week) {
  for (j in duration) {
    # i = week[1]
    # j = duration[1]
    
    data_treatment_control <- bioassay_data %>% 
      filter(Week == i) %>% 
      filter(Duration == j) %>% 
      group_by(Sample) %>% 
      mutate(sum_dead = sum(Dead)) %>% 
      mutate(sum_alive = sum(Alive)) %>% 
      select(Sample, Duration, sum_dead, sum_alive) %>% 
      distinct() %>% 
      select(Sample, sum_dead, sum_alive) %>% 
      arrange(Sample) %>%
      column_to_rownames(var = "Sample")
    
    if (nrow(data_treatment_control) > 0) {
      
      fisher.test(data_treatment_control, alternative = "less")
      
      one_sided_fisher_test_results_with_outliers <- fisher.test(data_treatment_control, alternative = "less")$p.value
      
      if (one_sided_fisher_test_results_with_outliers < 0.001) {
        one_sided_fisher_test_results_with_outliers <- "< 0.001"
      }
      
      assign(paste("one_sided_fisher_test_results_with_outliers_week", i, "duration", j, sep = "_"), one_sided_fisher_test_results_with_outliers)
      
      data_with_one_sided_fisher_test_results_with_outliers <- bioassay_data %>% 
        filter(Week == i) %>% 
        filter(Duration == j) %>% 
        group_by(Sample, Duration) %>%
        mutate(Average_Mortality = mean(`%Mortality`, na.rm = T)) %>% 
        select(Week, Sample, Duration, Average_Mortality) %>% 
        distinct() %>% 
        pivot_wider(id_cols = c(Duration), 
                    names_from = Sample, 
                    values_from = Average_Mortality, 
                    names_glue = "{Sample}_{.value}") %>%
        mutate(p.value = one_sided_fisher_test_results_with_outliers) %>% 
        mutate(`Control (% Mortality)` = `CONTROL_Average_Mortality`) %>% 
        mutate(`Treatment (% Mortality)` = `TREATMENT_Average_Mortality`) %>% 
        select(Duration, `Control (% Mortality)`, `Treatment (% Mortality)`, p.value)
      
      assign(paste("data_with_one_sided_fisher_test_results_with_outliers_week", i, "duration", j, sep = "_"), data_with_one_sided_fisher_test_results_with_outliers)
    }
  }
}
```

```{r}
#| echo: false
kable(bind_rows(data_with_one_sided_fisher_test_results_with_outliers_week_1_duration_24hr,
                data_with_one_sided_fisher_test_results_with_outliers_week_1_duration_48hr))
```

### Comparison of treatment to control (excluding outliers)

*Treatment replicate 1 from 48 hours was treated as an outlier and removed.* Mortality was significantly greater in the treatment than in the control for each experiment duration (24 and 48 hours) in week 1.

```{r}
#| echo: false

bioassay_data_control_outliers_week_1 <- bioassay_data_control_outliers %>% 
  filter(Week == 1)

bioassay_data_without_outliers_week_1 <- bind_rows(bioassay_data_control_outliers_week_1, bioassay_data_treatment_outliers_week_1) %>% 
  filter(Outlier == "No")
for (i in week) {
  for (j in duration) {
    i = week[1]
    # j = duration[1]
    
    data_treatment_control_without_outliers <- bioassay_data_without_outliers_week_1 %>% 
      
      filter(Duration == j) %>% 
      group_by(Sample) %>% 
      mutate(sum_dead = sum(Dead)) %>% 
      mutate(sum_alive = sum(Alive)) %>% 
      select(Sample, Duration, sum_dead, sum_alive) %>% 
      distinct() %>% 
      select(Sample, sum_dead, sum_alive) %>% 
      arrange(Sample) %>% 
      column_to_rownames(var = "Sample")
    
    if (nrow(data_treatment_control_without_outliers) > 0) {
      
      fisher.test(data_treatment_control_without_outliers, alternative = "less")
      
      one_sided_fisher_test_results_exclude_outliers <- fisher.test(data_treatment_control_without_outliers, alternative = "less")$p.value
      
      if (one_sided_fisher_test_results_exclude_outliers < 0.001) {
        one_sided_fisher_test_results_exclude_outliers <- "< 0.001"
        
      }
      
      assign(paste("one_sided_fisher_test_results_exclude_outliers_week", i, "duration", j, sep = "_"), one_sided_fisher_test_results_exclude_outliers)
      
      data_with_one_sided_fisher_test_results_exclude_outliers <- bioassay_data_without_outliers_week_1 %>% 
        filter(Duration == j) %>% 
        group_by(Sample, Duration) %>%
        mutate(Average_Mortality = mean(`%Mortality`, na.rm = T)) %>% 
        select(Week, Sample, Duration, Average_Mortality) %>% 
        distinct() %>% 
        pivot_wider(id_cols = c(Duration), 
                    names_from = Sample, 
                    values_from = Average_Mortality, 
                    names_glue = "{Sample}_{.value}") %>%
        mutate(p.value = one_sided_fisher_test_results_exclude_outliers) %>% 
        mutate(`Control (% Mortality)` = `CONTROL_Average_Mortality`) %>% 
        mutate(`Treatment (% Mortality)` = `TREATMENT_Average_Mortality`) %>% 
        mutate(`Control (% Mortality)` = round(`Control (% Mortality)`, digits = 1)) %>% 
        mutate(`Treatment (% Mortality)` = round(`Treatment (% Mortality)`, digits = 1)) %>%
        select(Duration, `Control (% Mortality)`, `Treatment (% Mortality)`, p.value)
      
      assign(paste("data_with_one_sided_fisher_test_results_exclude_outliers_week", i, "duration", j, sep = "_"), data_with_one_sided_fisher_test_results_exclude_outliers)
      
    }
  }
}
```

```{r}
#| echo: false

data_with_one_sided_fisher_test_results_exclude_outliers_week_1_all_durations <- bind_rows(
  data_with_one_sided_fisher_test_results_exclude_outliers_week_1_duration_24hr,
  data_with_one_sided_fisher_test_results_exclude_outliers_week_1_duration_48hr)

kable(data_with_one_sided_fisher_test_results_exclude_outliers_week_1_all_durations)

```

{{< pagebreak >}}

## Week 2 results

For week 2, there were four replicates (1, 2, 3, 4) for both the control and treatment groups. Mortality was assessed at 24, 48 and 72 hours.

### Mortality of control replicates

```{r}
#| include: false
bioassay_data %>% 
  filter(Week == "2") %>% 
  filter(Sample == "CONTROL") %>% 
  mutate(Outlier = if_else(`%Mortality` >= 10, "Yes", "No")) %>% 
  select(Duration, Replicate, `%Mortality`, Outlier)
```

During week 2, the following control replicates had mortality \>= 10%:

**Readings at 24 hours:** Control replicate 3 had 28% mortality. This replicate was treated as an outlier.

```{r}
#| echo: false
kable(bioassay_data %>% 
        filter(Week == "2") %>% 
        filter(Sample == "CONTROL") %>% 
        mutate(Outlier = if_else(`%Mortality` >= 10, "Yes", "No")) %>% 
        filter(Outlier == "Yes") %>% 
        select(Duration, Replicate, Dead, Total, `%Mortality`))
```

### Analysis of treatment replicates

There were no treatment replicates for week 2 that were significantly different from other treatment replicates.

```{r}
#| include: false

compare_all_trt_reps_pvalue_week_2_duration_24hr
compare_all_trt_reps_pvalue_week_2_duration_48hr
compare_all_trt_reps_pvalue_week_2_duration_72hr

kable(pairwise_fisher_test_results_week_2_duration_24hr)
kable(pairwise_fisher_test_results_week_2_duration_48hr)
kable(pairwise_fisher_test_results_week_2_duration_72hr)

```

```{r}
#| include: false
bioassay_data_treatment_outliers_week_2 <- bioassay_data %>% 
  filter(Sample == "TREATMENT") %>% 
  filter(Week == 2) %>% 
  mutate(Outlier = "No")

kable(bioassay_data_treatment_outliers_week_2)
```

### Comparison of treatment to control

Mortality was significantly greater in the treatment than in the control for each experiment duration (24, 48, and 72 hours) in week 2.

```{r}
#| echo: false
kable(bind_rows(data_with_one_sided_fisher_test_results_with_outliers_week_2_duration_24hr,
                data_with_one_sided_fisher_test_results_with_outliers_week_2_duration_48hr,
                data_with_one_sided_fisher_test_results_with_outliers_week_2_duration_72hr))
```

### Comparison of treatment to control (excluding outliers)

*Control replicate 3 from 24 hours was treated as an outlier and removed.* Mortality was significantly greater in the treatment than in the control for each experiment duration (24, 48, and 72 hours) in week 2.

```{r}
#| echo: false

bioassay_data_control_outliers_week_2 <- bioassay_data_control_outliers %>% 
  filter(Week == 2)

bioassay_data_without_outliers_week_2 <- bind_rows(bioassay_data_control_outliers_week_2, bioassay_data_treatment_outliers_week_2) %>% 
  filter(Outlier == "No")

for (i in week) {
  for (j in duration) {
    i = week[2]
    # j = duration[1]
    
    data_treatment_control_without_outliers <- bioassay_data_without_outliers_week_2 %>% 
      filter(Duration == j) %>% 
      group_by(Sample) %>% 
      mutate(sum_dead = sum(Dead)) %>% 
      mutate(sum_alive = sum(Alive)) %>% 
      select(Sample, Duration, sum_dead, sum_alive) %>% 
      distinct() %>% 
      select(Sample, sum_dead, sum_alive) %>% 
      arrange(Sample) %>% 
      column_to_rownames(var = "Sample")
    
    if (nrow(data_treatment_control_without_outliers) > 0) {
      
      fisher.test(data_treatment_control_without_outliers, alternative = "less")
      
      one_sided_fisher_test_results_exclude_outliers <- fisher.test(data_treatment_control_without_outliers, alternative = "less")$p.value
      
      if (one_sided_fisher_test_results_exclude_outliers < 0.001) {
        one_sided_fisher_test_results_exclude_outliers <- "< 0.001"
      }
      
      assign(paste("one_sided_fisher_test_results_exclude_outliers_week", i, "duration", j, sep = "_"), one_sided_fisher_test_results_exclude_outliers)
      
      data_with_one_sided_fisher_test_results_exclude_outliers <- bioassay_data_without_outliers_week_2 %>% 
        filter(Duration == j) %>% 
        group_by(Sample, Duration) %>%
        mutate(Average_Mortality = mean(`%Mortality`, na.rm = T)) %>% 
        select(Week, Sample, Duration, Average_Mortality) %>% 
        distinct() %>% 
        pivot_wider(id_cols = c(Duration), 
                    names_from = Sample, 
                    values_from = Average_Mortality, 
                    names_glue = "{Sample}_{.value}") %>%
        mutate(p.value = one_sided_fisher_test_results_exclude_outliers) %>% 
        mutate(`Control (% Mortality)` = `CONTROL_Average_Mortality`) %>% 
        mutate(`Treatment (% Mortality)` = `TREATMENT_Average_Mortality`) %>% 
        mutate(`Control (% Mortality)` = round(`Control (% Mortality)`, digits = 1)) %>% 
        mutate(`Treatment (% Mortality)` = round(`Treatment (% Mortality)`, digits = 1)) %>%
        select(Duration, `Control (% Mortality)`, `Treatment (% Mortality)`, p.value)
      
      assign(paste("data_with_one_sided_fisher_test_results_exclude_outliers_week", i, "duration", j, sep = "_"), data_with_one_sided_fisher_test_results_exclude_outliers)
      
    }
  }
}
```

```{r}
#| echo: false

data_with_one_sided_fisher_test_results_exclude_outliers_week_2_all_durations <- bind_rows(
  data_with_one_sided_fisher_test_results_exclude_outliers_week_2_duration_24hr,
  data_with_one_sided_fisher_test_results_exclude_outliers_week_2_duration_48hr,
  data_with_one_sided_fisher_test_results_exclude_outliers_week_2_duration_72hr
)
kable(data_with_one_sided_fisher_test_results_exclude_outliers_week_2_all_durations)

```

{{< pagebreak >}}

## Week 3 results

For week 3, there were four replicates (1, 2, 3, 4) for both the control and treatment groups. Mortality was assessed at 24, 48, 72, 96 hours.

### Mortality of control replicates

```{r}
#| include: false
bioassay_data %>% 
  filter(Week == "3") %>% 
  filter(Sample == "CONTROL") %>% 
  mutate(Outlier = if_else(`%Mortality` >= 10, "Yes", "No")) %>% 
  select(Duration, Replicate, `%Mortality`, Outlier)
```

During week 3, the following control replicates had mortality \>= 10%:

**Readings at 48 hours:** Control replicate 2 had 16% mortality. This replicate was treated as an outlier.

```{r}
#| echo: false
kable(bioassay_data %>% 
        filter(Week == "3") %>% 
        filter(Sample == "CONTROL") %>% 
        mutate(Outlier = if_else(`%Mortality` >= 10, "Yes", "No")) %>% 
        filter(Outlier == "Yes") %>%
        select(Duration, Replicate, Dead, Total, `%Mortality`))
```

### Analysis of treatment replicates

During week 3, readings taken at 48 and 96 hours had significantly different mortality between replicates.

**Readings at 48 hours:** Replicate 4 (64% mortality) was significantly different from replicate 1 (100% mortality) and replicate 2 (100% mortality). Replicate 4 was treated as an outlier.

**Readings at 96 hours:** Replicate 1 (64% mortality) was significantly different from replicate 2 (100% mortality), replicate 3 (96% mortality) and replicate 4 (100% mortality). Replicate 1 was treated as an outlier.

```{r}
#| include: false

compare_all_trt_reps_pvalue_week_3_duration_24hr
compare_all_trt_reps_pvalue_week_3_duration_48hr
compare_all_trt_reps_pvalue_week_3_duration_72hr
compare_all_trt_reps_pvalue_week_3_duration_96hr

kable(pairwise_fisher_test_results_week_3_duration_24hr)
kable(pairwise_fisher_test_results_week_3_duration_48hr)
kable(pairwise_fisher_test_results_week_3_duration_72hr)
kable(pairwise_fisher_test_results_week_3_duration_96hr)

```

```{r}
#| include: false
bioassay_data_treatment_outliers_week_3 <- bioassay_data %>% 
  filter(Sample == "TREATMENT") %>% 
  filter(Week == 3) %>% 
  mutate(Outlier = if_else(Duration %in% c("48hr") & Replicate == "4" | 
                             Duration %in% c("96hr") & Replicate == "1", 
                           "Yes", "No"))

kable(bioassay_data_treatment_outliers_week_3)
```

| Duration | Replicates for comparison | Adjusted p-value |
|----------|---------------------------|-----------------:|
| 48hr     | 1, 4                      |          0.00978 |
| 48hr     | 2, 4                      |          0.00978 |
| 96hr     | 1, 2                      |          0.00978 |
| 96hr     | 1, 3                      |          0.04240 |
| 96hr     | 1, 4                      |          0.00978 |

### Comparison of treatment to control

Mortality was significantly greater in the treatment than in the control for each experiment duration (24, 28, 72, and 96 hours) in week 3.

```{r}
#| echo: false
kable(bind_rows(data_with_one_sided_fisher_test_results_with_outliers_week_3_duration_24hr,
                data_with_one_sided_fisher_test_results_with_outliers_week_3_duration_48hr,
                data_with_one_sided_fisher_test_results_with_outliers_week_3_duration_72hr,
                data_with_one_sided_fisher_test_results_with_outliers_week_3_duration_96hr))
```

### Comparison of treatment to control (excluding outliers)

*Control replicate 2 from 48 hours was treated as an outlier and removed. Treatment replicate 4 from 48 hours and treatment replicate 1 from 96 hours were treated as outliers and removed.* Mortality was significantly greater in the treatment than in the control for each experiment duration (24, 28, 72, and 96 hours) in week 3.

```{r}
#| echo: false

bioassay_data_control_outliers_week_3 <- bioassay_data_control_outliers %>% 
  filter(Week == 3)

bioassay_data_without_outliers_week_3 <- bind_rows(bioassay_data_control_outliers_week_3, bioassay_data_treatment_outliers_week_3) %>% 
  filter(Outlier == "No")

for (i in week) {
  for (j in duration) {
    i = week[3]
    # j = duration[1]
    
    data_treatment_control_without_outliers <- bioassay_data_without_outliers_week_3 %>% 
      filter(Duration == j) %>% 
      group_by(Sample) %>% 
      mutate(sum_dead = sum(Dead)) %>% 
      mutate(sum_alive = sum(Alive)) %>% 
      select(Sample, Duration, sum_dead, sum_alive) %>% 
      distinct() %>% 
      select(Sample, sum_dead, sum_alive) %>% 
      arrange(Sample) %>% 
      column_to_rownames(var = "Sample")
    
    if (nrow(data_treatment_control_without_outliers) > 0) {
      
      fisher.test(data_treatment_control_without_outliers, alternative = "less")
      
      one_sided_fisher_test_results_exclude_outliers <- fisher.test(data_treatment_control_without_outliers, alternative = "less")$p.value
      
      if (one_sided_fisher_test_results_exclude_outliers < 0.001) {
        one_sided_fisher_test_results_exclude_outliers <- "< 0.001"
      }
      
      assign(paste("one_sided_fisher_test_results_exclude_outliers_week", i, "duration", j, sep = "_"), one_sided_fisher_test_results_exclude_outliers)
      
      data_with_one_sided_fisher_test_results_exclude_outliers <- bioassay_data_without_outliers_week_3 %>% 
        filter(Duration == j) %>% 
        group_by(Sample, Duration) %>%
        mutate(Average_Mortality = mean(`%Mortality`, na.rm = T)) %>% 
        select(Week, Sample, Duration, Average_Mortality) %>% 
        distinct() %>% 
        pivot_wider(id_cols = c(Duration), 
                    names_from = Sample, 
                    values_from = Average_Mortality, 
                    names_glue = "{Sample}_{.value}") %>%
        mutate(p.value = one_sided_fisher_test_results_exclude_outliers) %>% 
        mutate(`Control (% Mortality)` = `CONTROL_Average_Mortality`) %>% 
        mutate(`Treatment (% Mortality)` = `TREATMENT_Average_Mortality`) %>% 
        mutate(`Control (% Mortality)` = round(`Control (% Mortality)`, digits = 1)) %>% 
        mutate(`Treatment (% Mortality)` = round(`Treatment (% Mortality)`, digits = 1)) %>%
        select(Duration, `Control (% Mortality)`, `Treatment (% Mortality)`, p.value)
      
      assign(paste("data_with_one_sided_fisher_test_results_exclude_outliers_week", i, "duration", j, sep = "_"), data_with_one_sided_fisher_test_results_exclude_outliers)
      
    }
  }
}
```

```{r}
#| echo: false

data_with_one_sided_fisher_test_results_exclude_outliers_week_3_all_durations <- bind_rows(
  data_with_one_sided_fisher_test_results_exclude_outliers_week_3_duration_24hr,
  data_with_one_sided_fisher_test_results_exclude_outliers_week_3_duration_48hr,
  data_with_one_sided_fisher_test_results_exclude_outliers_week_3_duration_72hr,
  data_with_one_sided_fisher_test_results_exclude_outliers_week_3_duration_96hr
)
kable(data_with_one_sided_fisher_test_results_exclude_outliers_week_3_all_durations)

```

{{< pagebreak >}}

## Week 4 results

For week 4, there were four replicates (1, 2, 3, 4) for both the control and treatment groups. Mortality was assessed at 24, 48, 72, and 96 hours.

### Mortality of control replicates

```{r}
#| include: false
bioassay_data %>% 
  filter(Week == "4") %>% 
  filter(Sample == "CONTROL") %>% 
  mutate(Outlier = if_else(`%Mortality` >= 10, "Yes", "No")) %>% 
  select(Duration, Replicate, `%Mortality`, Outlier)
```

During week 4, the following control replicates had mortality \>= 10%:

**Readings at 24 hours:** Control replicate 2 had 52% mortality. This replicate was treated as an outlier.

```{r}
#| echo: false
kable(bioassay_data %>% 
        filter(Week == "4") %>% 
        filter(Sample == "CONTROL") %>% 
        mutate(Outlier = if_else(`%Mortality` >= 10, "Yes", "No")) %>% 
        filter(Outlier == "Yes") %>% 
        select(Duration, Replicate, Dead, Total, `%Mortality`))
```

### Analysis of treatment replicates

```{r}
#| include: false

compare_all_trt_reps_pvalue_week_4_duration_24hr
compare_all_trt_reps_pvalue_week_4_duration_48hr
compare_all_trt_reps_pvalue_week_4_duration_72hr
compare_all_trt_reps_pvalue_week_4_duration_96hr

kable(pairwise_fisher_test_results_week_4_duration_24hr)
kable(pairwise_fisher_test_results_week_4_duration_48hr)
kable(pairwise_fisher_test_results_week_4_duration_72hr)
kable(pairwise_fisher_test_results_week_4_duration_96hr)

```

During week 4, readings taken at 48 hours had significantly different mortality between replicates.

**Readings at 48 hours:** Replicate 2 (68% mortality) was significantly different from replicate 1 (100% mortality), replicate 3 (100% mortality), and replicate 4 (100% mortality). Replicate 2 was treated as an outlier.

```{r}
#| include: false
bioassay_data_treatment_outliers_week_4 <- bioassay_data %>% 
  filter(Sample == "TREATMENT") %>% 
  filter(Week == 4) %>%
  mutate(Outlier = if_else(Duration %in% c("48hr") & Replicate == "2", "Yes", "No"))

kable(bioassay_data_treatment_outliers_week_4)

```

| Duration | Replicates for comparison | Adjusted p-value |
|----------|---------------------------|-----------------:|
| 48hr     | 1, 2                      |           0.0242 |
| 48hr     | 2, 3                      |           0.0242 |
| 48hr     | 2, 4                      |           0.0242 |

### Comparison of treatment to control

Mortality was significantly greater in the treatment than in the control for each experiment duration (24, 48, 72, and 96 hours) in week 4.

```{r}
#| echo: false
kable(bind_rows(data_with_one_sided_fisher_test_results_with_outliers_week_4_duration_24hr,
                data_with_one_sided_fisher_test_results_with_outliers_week_4_duration_48hr,
                data_with_one_sided_fisher_test_results_with_outliers_week_4_duration_72hr, 
                data_with_one_sided_fisher_test_results_with_outliers_week_4_duration_96hr))
```

### Comparison of treatment to control (excluding outliers)

*Control replicate 2 from 24 hours was treated as an outlier and removed. Treatment replicate 2 from 48 hours was treated as an outlier and removed.* Mortality was significantly greater in the treatment than in the control for each experiment duration (24, 48, 72, and 96 hours) in week 4.

```{r}
#| echo: false

bioassay_data_control_outliers_week_4 <- bioassay_data_control_outliers %>% 
  filter(Week == 4)

bioassay_data_without_outliers_week_4 <- bind_rows(bioassay_data_control_outliers_week_4, bioassay_data_treatment_outliers_week_4) %>% 
  filter(Outlier == "No")

for (i in week) {
  for (j in duration) {
    i = week[4]
    # j = duration[1]
    
    data_treatment_control_without_outliers <- bioassay_data_without_outliers_week_4 %>% 
      
      filter(Duration == j) %>% 
      group_by(Sample) %>% 
      mutate(sum_dead = sum(Dead)) %>% 
      mutate(sum_alive = sum(Alive)) %>% 
      select(Sample, Duration, sum_dead, sum_alive) %>% 
      distinct() %>% 
      select(Sample, sum_dead, sum_alive) %>% 
      arrange(Sample) %>% 
      column_to_rownames(var = "Sample")
    
    if (nrow(data_treatment_control_without_outliers) > 0) {
      
      fisher.test(data_treatment_control_without_outliers, alternative = "less")
      
      one_sided_fisher_test_results_exclude_outliers <- fisher.test(data_treatment_control_without_outliers, alternative = "less")$p.value
      
      if (one_sided_fisher_test_results_exclude_outliers < 0.001) {
        one_sided_fisher_test_results_exclude_outliers <- "< 0.001"
        
      }
      
      assign(paste("one_sided_fisher_test_results_exclude_outliers_week", i, "duration", j, sep = "_"), one_sided_fisher_test_results_exclude_outliers)
      
      data_with_one_sided_fisher_test_results_exclude_outliers <- bioassay_data_without_outliers_week_4 %>% 
        filter(Duration == j) %>% 
        group_by(Sample, Duration) %>%
        mutate(Average_Mortality = mean(`%Mortality`, na.rm = T)) %>% 
        select(Week, Sample, Duration, Average_Mortality) %>% 
        distinct() %>% 
        pivot_wider(id_cols = c(Duration), 
                    names_from = Sample, 
                    values_from = Average_Mortality, 
                    names_glue = "{Sample}_{.value}") %>%
        mutate(p.value = one_sided_fisher_test_results_exclude_outliers) %>% 
        mutate(`Control (% Mortality)` = `CONTROL_Average_Mortality`) %>% 
        mutate(`Treatment (% Mortality)` = `TREATMENT_Average_Mortality`) %>% 
        mutate(`Control (% Mortality)` = round(`Control (% Mortality)`, digits = 1)) %>% 
        mutate(`Treatment (% Mortality)` = round(`Treatment (% Mortality)`, digits = 1)) %>%
        select(Duration, `Control (% Mortality)`, `Treatment (% Mortality)`, p.value)
      
      assign(paste("data_with_one_sided_fisher_test_results_exclude_outliers_week", i, "duration", j, sep = "_"), data_with_one_sided_fisher_test_results_exclude_outliers)
      
    }
  }
}
```

```{r}
#| echo: false

data_with_one_sided_fisher_test_results_exclude_outliers_week_4_all_durations <- bind_rows(
  data_with_one_sided_fisher_test_results_exclude_outliers_week_4_duration_24hr,
  data_with_one_sided_fisher_test_results_exclude_outliers_week_4_duration_48hr,
  data_with_one_sided_fisher_test_results_exclude_outliers_week_4_duration_72hr,
  data_with_one_sided_fisher_test_results_exclude_outliers_week_4_duration_96hr
)

kable(data_with_one_sided_fisher_test_results_exclude_outliers_week_4_all_durations)

```

{{< pagebreak >}}

## Week 5 results

For week 5, there were four replicates (1, 2, 3, 4) for both the control and treatment groups. Mortality was assessed at 24 and 48 hours.

### Mortality of control replicates

```{r}
#| include: false
bioassay_data %>% 
  filter(Week == "5") %>% 
  filter(Sample == "CONTROL") %>% 
  mutate(Outlier = if_else(`%Mortality` >= 10, "Yes", "No")) %>% 
  select(Duration, Replicate, `%Mortality`, Outlier)
```

During week 5, the following control replicates had mortality \>= 10%:

**Readings at 24 hours:** Control replicate 1 had 12% mortality. This replicate was treated as an outlier.\
**Readings at 48 hours:** Control replicate 4 had 12% mortality. This replicate was treated as an outlier.

```{r}
#| echo: false
kable(bioassay_data %>% 
        filter(Week == "5") %>% 
        filter(Sample == "CONTROL") %>% 
        mutate(Outlier = if_else(`%Mortality` >= 10, "Yes", "No")) %>% 
        filter(Outlier == "Yes") %>% 
        select(Duration, Replicate, Dead, Total, `%Mortality`))
```

### Analysis of treatment replicates

```{r}
#| include: false

compare_all_trt_reps_pvalue_week_5_duration_24hr
compare_all_trt_reps_pvalue_week_5_duration_48hr

kable(pairwise_fisher_test_results_week_5_duration_24hr)
kable(pairwise_fisher_test_results_week_5_duration_48hr)

```

There were no treatment replicates for week 5 that were significantly different from other treatment replicates.

```{r}
#| include: false
bioassay_data_treatment_outliers_week_5 <- bioassay_data %>% 
  filter(Sample == "TREATMENT") %>% 
  filter(Week == 5) %>% 
  mutate(Outlier = "No")

kable(bioassay_data_treatment_outliers_week_5)
```

### Comparison of treatment to control

Mortality was significantly greater in the treatment than in the control for each experiment duration (24 and 48 hours) in week 5.

```{r}
#| echo: false
kable(bind_rows(data_with_one_sided_fisher_test_results_with_outliers_week_5_duration_24hr,
                data_with_one_sided_fisher_test_results_with_outliers_week_5_duration_48hr))
```

### Comparison of treatment to control (excluding outliers)

*Control replicate 1 from 24 hours and control replicate 4 from 48 hours were treated as outliers and removed.* Mortality was significantly greater in the treatment than in the control for each experiment duration (24 and 48 hours) in week 5.

```{r}
#| echo: false

bioassay_data_control_outliers_week_5 <- bioassay_data_control_outliers %>% 
  filter(Week == 5)

bioassay_data_without_outliers_week_5 <- bind_rows(bioassay_data_control_outliers_week_5, bioassay_data_treatment_outliers_week_5) %>% 
  filter(Outlier == "No")

for (i in week) {
  for (j in duration) {
    i = week[5]
    # j = duration[1]
    
    data_treatment_control_without_outliers <- bioassay_data_without_outliers_week_5 %>% 
      
      filter(Duration == j) %>% 
      group_by(Sample) %>% 
      mutate(sum_dead = sum(Dead)) %>% 
      mutate(sum_alive = sum(Alive)) %>% 
      select(Sample, Duration, sum_dead, sum_alive) %>% 
      distinct() %>% 
      select(Sample, sum_dead, sum_alive) %>% 
      arrange(Sample) %>% 
      column_to_rownames(var = "Sample")
    
    if (nrow(data_treatment_control_without_outliers) > 0) {
      
      fisher.test(data_treatment_control_without_outliers, alternative = "less")
      
      one_sided_fisher_test_results_exclude_outliers <- fisher.test(data_treatment_control_without_outliers, alternative = "less")$p.value
      
      if (one_sided_fisher_test_results_exclude_outliers < 0.001) {
        one_sided_fisher_test_results_exclude_outliers <- "< 0.001"
      }
      
      assign(paste("one_sided_fisher_test_results_exclude_outliers_week", i, "duration", j, sep = "_"), one_sided_fisher_test_results_exclude_outliers)
      
      data_with_one_sided_fisher_test_results_exclude_outliers <- bioassay_data_without_outliers_week_5 %>% 
        filter(Duration == j) %>% 
        group_by(Sample, Duration) %>%
        mutate(Average_Mortality = mean(`%Mortality`, na.rm = T)) %>% 
        select(Week, Sample, Duration, Average_Mortality) %>% 
        distinct() %>% 
        pivot_wider(id_cols = c(Duration), 
                    names_from = Sample, 
                    values_from = Average_Mortality, 
                    names_glue = "{Sample}_{.value}") %>%
        mutate(p.value = one_sided_fisher_test_results_exclude_outliers) %>% 
        mutate(`Control (% Mortality)` = `CONTROL_Average_Mortality`) %>% 
        mutate(`Treatment (% Mortality)` = `TREATMENT_Average_Mortality`) %>% 
        mutate(`Control (% Mortality)` = round(`Control (% Mortality)`, digits = 1)) %>% 
        mutate(`Treatment (% Mortality)` = round(`Treatment (% Mortality)`, digits = 1)) %>%
        select(Duration, `Control (% Mortality)`, `Treatment (% Mortality)`, p.value)
      
      assign(paste("data_with_one_sided_fisher_test_results_exclude_outliers_week", i, "duration", j, sep = "_"), data_with_one_sided_fisher_test_results_exclude_outliers)
      
    }
  }
}
```

```{r}
#| echo: false

data_with_one_sided_fisher_test_results_exclude_outliers_week_5_all_durations <- bind_rows(
  data_with_one_sided_fisher_test_results_exclude_outliers_week_5_duration_24hr,
  data_with_one_sided_fisher_test_results_exclude_outliers_week_5_duration_48hr
)
kable(data_with_one_sided_fisher_test_results_exclude_outliers_week_5_all_durations)

```

{{< pagebreak >}}

## Week 6 results

For week 6, there were four replicates (1, 2, 3, 4) for both the control and treatment groups. Mortality was assessed at 24, 48, and 72 hours.

### Mortality of control replicates

```{r}
#| include: false
bioassay_data %>% 
  filter(Week == "6") %>% 
  filter(Sample == "CONTROL") %>% 
  mutate(Outlier = if_else(`%Mortality` >= 10, "Yes", "No")) %>% 
  select(Duration, Replicate, `%Mortality`, Outlier)
```

During week 6, the following control replicates had mortality \>= 10%:

**Readings at 72 hours:** Control replicate 3 had 16% mortality, and control replicate 4 had 12% mortality. These replicates were treated as outliers.

```{r}
#| echo: false
kable(bioassay_data %>% 
        filter(Week == "6") %>% 
        filter(Sample == "CONTROL") %>% 
        mutate(Outlier = if_else(`%Mortality` >= 10, "Yes", "No")) %>% 
        filter(Outlier == "Yes") %>% 
        select(Duration, Replicate, Dead, Total, `%Mortality`))
```

### Analysis of treatment replicates

```{r}
#| include: false

compare_all_trt_reps_pvalue_week_6_duration_24hr
compare_all_trt_reps_pvalue_week_6_duration_48hr
compare_all_trt_reps_pvalue_week_6_duration_72hr

kable(pairwise_fisher_test_results_week_6_duration_24hr)
kable(pairwise_fisher_test_results_week_6_duration_48hr)
kable(pairwise_fisher_test_results_week_6_duration_72hr)

```

There were no treatment replicates for week 6 that were significantly different from other treatment replicates.

```{r}
#| include: false
bioassay_data_treatment_outliers_week_6 <- bioassay_data %>% 
  filter(Sample == "TREATMENT") %>% 
  filter(Week == 6) %>% 
  mutate(Outlier = "No")

kable(bioassay_data_treatment_outliers_week_6)
```

### Comparison of treatment to control

Mortality was significantly greater in the treatment than in the control for each experiment duration (24, 48, and 72 hours) in week 6.

```{r}
#| echo: false
kable(bind_rows(data_with_one_sided_fisher_test_results_with_outliers_week_6_duration_24hr,
                data_with_one_sided_fisher_test_results_with_outliers_week_6_duration_48hr,
                data_with_one_sided_fisher_test_results_with_outliers_week_6_duration_72hr))
```

### Comparison of treatment to control (excluding outliers)

*Control replicates 3 and 4 from 72 hours were treated as outliers and removed.* Mortality was significantly greater in the treatment than in the control for each experiment duration (24, 48, and 72 hours) in week 6.

```{r}
#| echo: false

bioassay_data_control_outliers_week_6 <- bioassay_data_control_outliers %>% 
  filter(Week == 6)

bioassay_data_without_outliers_week_6 <- bind_rows(bioassay_data_control_outliers_week_6, bioassay_data_treatment_outliers_week_6) %>% 
  filter(Outlier == "No")

for (i in week) {
  for (j in duration) {
    i = week[6]
    # j = duration[1]
    
    data_treatment_control_without_outliers <- bioassay_data_without_outliers_week_6 %>% 
      filter(Duration == j) %>% 
      group_by(Sample) %>% 
      mutate(sum_dead = sum(Dead)) %>% 
      mutate(sum_alive = sum(Alive)) %>% 
      select(Sample, Duration, sum_dead, sum_alive) %>% 
      distinct() %>% 
      select(Sample, sum_dead, sum_alive) %>% 
      arrange(Sample) %>% 
      column_to_rownames(var = "Sample")
    
    if (nrow(data_treatment_control_without_outliers) > 0) {
      
      fisher.test(data_treatment_control_without_outliers, alternative = "less")
      
      one_sided_fisher_test_results_exclude_outliers <- fisher.test(data_treatment_control_without_outliers, alternative = "less")$p.value
      
      if (one_sided_fisher_test_results_exclude_outliers < 0.001) {
        one_sided_fisher_test_results_exclude_outliers <- "< 0.001"
      }
      
      assign(paste("one_sided_fisher_test_results_exclude_outliers_week", i, "duration", j, sep = "_"), one_sided_fisher_test_results_exclude_outliers)
      
      data_with_one_sided_fisher_test_results_exclude_outliers <- bioassay_data_without_outliers_week_6 %>% 
        filter(Duration == j) %>% 
        group_by(Sample, Duration) %>%
        mutate(Average_Mortality = mean(`%Mortality`, na.rm = T)) %>% 
        select(Week, Sample, Duration, Average_Mortality) %>% 
        distinct() %>% 
        pivot_wider(id_cols = c(Duration), 
                    names_from = Sample, 
                    values_from = Average_Mortality, 
                    names_glue = "{Sample}_{.value}") %>%
        mutate(p.value = one_sided_fisher_test_results_exclude_outliers) %>% 
        mutate(`Control (% Mortality)` = `CONTROL_Average_Mortality`) %>% 
        mutate(`Treatment (% Mortality)` = `TREATMENT_Average_Mortality`) %>% 
        mutate(`Control (% Mortality)` = round(`Control (% Mortality)`, digits = 1)) %>% 
        mutate(`Treatment (% Mortality)` = round(`Treatment (% Mortality)`, digits = 1)) %>%
        select(Duration, `Control (% Mortality)`, `Treatment (% Mortality)`, p.value)
      
      assign(paste("data_with_one_sided_fisher_test_results_exclude_outliers_week", i, "duration", j, sep = "_"), data_with_one_sided_fisher_test_results_exclude_outliers)
      
    }
  }
}
```

```{r}
#| echo: false

data_with_one_sided_fisher_test_results_exclude_outliers_week_6_all_durations <- bind_rows(
  data_with_one_sided_fisher_test_results_exclude_outliers_week_6_duration_24hr,
  data_with_one_sided_fisher_test_results_exclude_outliers_week_6_duration_48hr,
  data_with_one_sided_fisher_test_results_exclude_outliers_week_6_duration_72hr
)
kable(data_with_one_sided_fisher_test_results_exclude_outliers_week_6_all_durations)

```
